Analysis plan

### 2023-02-06 02. Design/MSG-P2312-Analysis Plan-1-clusters.rmd
Update MSG-P2312-Analysis Plan-1-clusters.rmd
Adding more details to the study design and descriptive statistics

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(janitor)
library(skimr)
library(DataExplorer)
library(readr)
library(oescolorrrs)
library(oesrrr)
library(flextable)
```

@@

### Table 1. GSA Daily Check-In Data Variables Available Feb 2022-Dec 2022.

@@

### Table 2. GSA Roster Variables.

@@

We have no plans to exclude any date.

@@

To determine the number appropriate number of clusters, several methods
exist. A rule of thumb to identify the correct number us to use

@@

@@

@@

Proportion in staff time telework

Proportion in staff time telework by the day of the week

Proportion in staff time onsite

Proportion in staff time onsite by the day of the week

Correlations of by staff time between the days of the week

Size of teams

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(janitor)
library(skimr)
library(DataExplorer)
library(readr)
library(oescolorrrs)
library(oesrrr)
library(flextable)
library(tidymodels)
library(ggthemes)
```

@@

#### Table 1. GSA Daily Check-In Data Variables Available Feb 2022-Dec 2022.

@@

#### Table 2. GSA Roster Variables.

@@

We have no plans to exclude any data.

@@

To determine the number appropriate number of clusters, several methods
exist. A rule of thumb to identify the correct number is to use the
number calculated from:

@@

However, for this work, the *elbow method* will be used. To do this, the
k-means algorithm will be executed for different numbers of clusters
from 1 to 20. For each number of clusters, a total within sum of squares
will be calculated. As the number of number of clusters increases the
total within sum of squares will decrease. The optimal number of
clusters is when the decrease in total within sum of squares starts
tapering off.

The figure below shows an example of the this. The total within sum of
squares decreases with the number of clusters. In this example, three
clusters appears to be the point where the returns to more cluster
diminishes.

#### Figure 1. Elbow method example

```{r}
set.seed(27)

centers <- tibble(
  cluster = factor(1:3), 
  num_points = c(100, 150, 50),  # number points in each cluster
  x1 = c(5, 0, -3),              # x1 coordinate of cluster center
  x2 = c(-1, 1, -2)              # x2 coordinate of cluster center
)

labelled_points <- 
  centers %>%
  mutate(
    x1 = map2(num_points, x1, rnorm),
    x2 = map2(num_points, x2, rnorm)
  ) %>% 
  select(-num_points) %>% 
  unnest(cols = c(x1, x2))

points <- 
  labelled_points %>% 
  select(-cluster)

kclusts <- 
  tibble(k = 1:20) %>%
  mutate(
    kclust = map(k, ~kmeans(points, .x)),
    tidied = map(kclust, tidy),
    glanced = map(kclust, glance),
    augmented = map(kclust, augment, points)
  )

clusters <- 
  kclusts %>%
  unnest(cols = c(tidied))

assignments <- 
  kclusts %>% 
  unnest(cols = c(augmented))

clusterings <- 
  kclusts %>%
  unnest(cols = c(glanced))

ggplot(clusterings, aes(k, tot.withinss)) +
  geom_line() +
  geom_point() +
  theme_minimal()

```

After the optimal number of clusters is determine, the clusters from
that iteration will be used to categorized the staff and team occupancy.

The work pattern statistics described below will be used to
characterized the clusters defined by the k-means algorithm.

@@

This is descriptive work and we will not be making any causal claims.

@@

The principle work pattern statistics are as follows:

-   Staff work pattern statistics:

    -   Proportion in staff time telework

    -   Proportion in staff time telework by the day of the week

    -   Proportion in staff time onsite

    -   Proportion in staff time onsite by the day of the week

    -   Correlations of by staff time between the days of the week

    -   Size of teams

    -   Number exiting employment

    -   Number changing supervisors

-   Team work pattern statistics:

    -   Proportion in staff time telework

    -   Proportion in staff time onsite

    -   Proportion of concurrent remote days for team members

    -   Proportion of concurrent onsite days for team members

    -   Size of teams

The descriptive statistics are:

-   Staff descriptive statistics

    -   Number and proportion of staff by grade

    -   Number and proportion of staff by occupational series

    -   Number and proportion of staff by position category

    -   Number and proportion of staff by GSA region

    -   Number and proportion of staff by SSO

    -   Number and proportion of staff by state

    -   Number and proportion of staff by city

    -   Number and proportion of staff by supervisor level

-   Team descriptive statistics

    -   Number and proportion of staff by most frequent grade

    -   Number and proportion of staff by most frequent occupational series

    -   Number and proportion of staff by most frequent position category

    -   Number and proportion of staff by most frequent GSA region

    -   Number and proportion of staff by most frequent SSO

    -   Number and proportion of staff by most frequent state

    -   Number and proportion of staff by most frequent city

    -   Number and proportion of staff by most frequent supervisor level
